---
title: "Mini Project 3"
author: "Sam Ellinwood, Haley McAtee, Sam Schneider"
date: "11/21/2017"
output: html_document
---

##Problem Statement

We want to determine if the movie industry has recently developed a higher propensity for generating sequels, remakes, and reboots. To evaluate this question, we plot the total number of these movies in a given year over time. Additionally, we examine how the number of years between a sequel and its original has changed over time. 


```{r, message = FALSE, warning = FALSE}
#Load packages
library(tidyverse)
library(mdsr)
library(RMySQL)
library(viridis)
db <- dbConnect_scidb(dbname = "imdb")
```

```{r}
#Create a table of sequels from SQL

sql_tbl <- db %>%
  dbGetQuery(
  #Select title and production year of sequels and linking variables from the aka_title and  movie_link tables in the IMDb dataset
  #Join the aka_title and movie_link tables
  #Filter out incomplete rows in the new table
  #Use kind_id = 1 to filter out any IMDB data not related to movies
  #Use link_type_id = 1 to filter out any movies that are not sequels
    "SELECT DISTINCT movie_link.movie_id AS sequel_id, linked_movie_id AS original_id, link_type_id, production_year AS sequel_year, title AS sequel_title  
  FROM imdb.movie_link
  JOIN aka_title ON movie_link.movie_id = aka_title.movie_id
  WHERE movie_link.movie_id > 0
    AND aka_title.kind_id = 1
    AND (link_type_id = 1);
")
```

```{r}
#Create a table of all movies with linking variables  (movies linked to sequels)
orig_tbl <- db %>%
  dbGetQuery(
    #Select title, production year, and linking variables from the aka_title and movie_link tables in the IMDb dataset
    #Join aka_title and movie_link tables
    #filter out rows that have a production year of NULL and/or movie_id of 0
    "SELECT movie_id AS original_id, production_year AS original_year, title AS original_title
    FROM imdb.aka_title
    WHERE production_year IS NOT NULL
     AND movie_id > 0;
")
```

```{r}
#Create the full table
full_tbl <- sql_tbl %>%
  #Use a left join to join the sequel tables and the table of all movies using original_id
  left_join(orig_tbl, by = "original_id") %>%
  #Remove duplicate rows
  unique() 
```


```{r}
#Create the data graphic
full_tbl %>%
  filter(sequel_year > 1950 & sequel_year < 2017 & original_year != "NA") %>%
  #Group by year
  group_by(sequel_year) %>%
  #Create two new aggregate variables. Count is the total number of sequels for each year. Difference is the average difference of years between    sequels and original films in each year
  summarize("count" = n(),
            "difference" = mean(sequel_year - original_year)) %>%
ggplot(aes(x = sequel_year, y = count)) +
  geom_point(aes(color = difference), size = 4) +
  geom_smooth(size = 2, se = FALSE) +
  scale_color_viridis(option = "inferno") +
  labs(x = "Year",
       y = "Number of Sequels", 
       color = "Average number of \nyears between\nsequel and original",
       title = "Number of Movie Sequels Over Time") +
  scale_x_continuous(breaks = seq(1950, 2017, 10)) +
  scale_y_continuous(breaks = seq(1000, 8000, 2000)) +
  theme(panel.grid = element_blank())
```


This data graphic shows that the number of sequels made has steadily increased over time, and the average number of years between the sequel release date and original movie release date has also increased over time. This leads us to believe that movie-goers increasingly demand sequels compared to in the past.

##Interpretation and Further Research
Alternative theories exist that explain this upward trend of the number of sequels produced. Analysts have found that the most common sequels are those that follow successful originals. Since box office revenues drive the success of movies, producers have been increasingly eager to capitalize on the original's success by producing a sequel that will be well known by the average movie-goer, which is less-risky than producing a new idea (Snow, 2016). Secondly, more and more sequels are following the "Saga" track in which these sequels build upon already successful storylines. Snow (2016) points out that the only two sequels to ever have won Oscars, *The Godfather Part II* and *Return of the King*, were both Sagas. Snow (2016) writes that "The paradox of the movie business is that while original movies change the game, safe bets pay the players." Producers have increasingly turned to remaking movies as a safe way to ensure high profits.

Although this graphic shows that there is a trend of increasing sequels over time, this does not necessarily mean that for a given movie, a sequel is more likely to be produced today than it was in 1950. Since there are more movies produced in total today than in 1950, there are more sequels produced. Thus, further study could compare the number of sequels produced as a percentage of the total number of movies produced in each year to more accurately depict the trend and compare it across different time periods. 

###GitHub Link
https://github.com/sellinwood8/sds192_mp3

###External Research Source
https://www.linkedin.com/pulse/why-hollywood-makes-more-sequels-every-year-even-though-shane-snow/
